---
output:
  pdf_document: default
  html_document: default
---
# Bellabeat Analysis in R
### David Jenkins
### Sep 30, 2021

Bellabeat is a manufacturer of high-tech, health-based products for women.
We have Urška Sršen who is the co-founder and chief creative officer,  Sando Mur who is also a co-founder and mathematician for the company.
Along with them we have the marketing analytics team, who is responsible for collecting, analyzing and reporting data that helps with marketing the products.
Bellabeat offers an app, a wellness tracker, a watch and a water bottle that tracks daily water intake that communicates with the Bellabeat app.
Bellabeat also offers a premium subscription for personalized guidance on various lifestyle categories.
 A link to their website can be found here. <https://bellabeat.com>
 
### Loading the packages needed for analysis
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
```
 
### Loading the data for analysis
```{r}
read.csv("C:/Users/jenki/OneDrive/Documents/dailyActivity_merged.csv")->daily_activity
read.csv("C:/Users/jenki/OneDrive/sleepDay_merged.csv")->sleep_data
read.csv("C:/Users/jenki/OneDrive/weightLogInfo_merged.csv")->weight
read.csv("C:/Users/jenki/OneDrive/hourlyCalories_merged.csv")->hour_cal
```
These datasets are available on Kaggle.com as part of public domain(FitBit Fitness Tracker Data provided by 'Mobius').
<https://www.kaggle.com/arashnic/fitbit/activity>.
I loaded the datasets into Excel and checked their validity there. I couldn't find any data for water intake.

## *How can we get users to use the tracker and the app more?
## *How can we highlight other important areas of the app?
## *How could these insights help with Bellabeat marketing strategy?

### Let's take a quick look at the data
```{r}
head(daily_activity)
head(sleep_data)
head(weight)
head(hour_cal)
```

### Cleaning the data
```{r}
rename(sleep_data, Date="SleepDay")->sleep_data
rename(daily_activity, Date="ActivityDate")->daily_activity
rename(hour_cal, Date="ActivityHour")->hour_cal
weight%>%separate(col=Date, into = c("Date","Time"), sep=" ")->weight
sleep_data%>%separate(col=Date, into = c("Date", "Time"), sep = " ")->sleep_data
hour_cal%>%separate(col=Date, into = c("Date", "Time"), sep=" ")->hour_cal
```

### How many participants do we have for each dataset?
```{r}
n_distinct(daily_activity$Id)
n_distinct(sleep_data$Id)
n_distinct(weight$Id)
n_distinct(hour_cal$Id)
```
*Daily Activity and Hourly Calorie has 33 participants.
*Sleep Day data has 24.
*Weight data has 8. Since weight data has so few participants, we can stress the importance of keeping up with your weight. The dataset will have no use for analysis.

### Merging tables for analysis
```{r}
merge(sleep_data, daily_activity, by=c("Id", "Date"))->combined_data
head(combined_data)
str(combined_data)
```

### Summary of combined data
```{r}
combined_data%>%
  select(TotalSteps, TrackerDistance,VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes, TotalSleepRecords, TotalTimeInBed, TotalMinutesAsleep, Calories)%>%
  summary()
```
#### *Average total steps among users is 8541.
#### *Average tracker distance is 6.034.
#### *Average calories burned is 2398.
#### *Average hourly calories burned is 97.39.
#### *Average sleep time is 419.5 minutes or around 6 hours with an average of 1.2 sleep records per user.
#### *Very active minutes and fairly active minutes are very low, in my opinion. Adding hourly motivators can help with motivating users to get active.

### Comparing total steps and active minutes
```{r}
ggplot(data=combined_data, aes(x=TotalSteps, y=VeryActiveMinutes)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "Total Steps vs Very Active Minutes")+theme(plot.title=element_text(hjust=0.5))
ggplot(data=combined_data, aes(x=TotalSteps, y=FairlyActiveMinutes)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "Total Steps vs Fairly Active Minutes")+theme(plot.title=element_text(hjust=0.5))
ggplot(data=combined_data, aes(x=TotalSteps, y=LightlyActiveMinutes)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "Total Steps vs Lightly Active Minutes")+theme(plot.title=element_text(hjust=0.5))
ggplot(data=combined_data, aes(x=TotalSteps, y=SedentaryMinutes)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "Total Steps vs Sedentary Minutes")+theme(plot.title=element_text(hjust=0.5))
```

#### The more steps you take, the more active you are according to the data.

### Comparing total step and calories burned.
```{r}
ggplot(data=combined_data, aes(x=TotalSteps, y=Calories)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "TotalSteps vs Calories")+theme(plot.title=element_text(hjust=0.5))
```

#### As we can see from this plot, the more steps are taken, the more calories are burned.

### Comparing sleep data.
```{r}
ggplot(data=combined_data, aes(x=TotalSteps, y=TotalMinutesAsleep)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "TotalSteps vs Total Minutes Asleep")+theme(plot.title=element_text(hjust=0.5))
ggplot(data=combined_data, aes(x=TotalSteps, y=TotalTimeInBed)) + geom_point(color="NavyBlue")+geom_smooth(color="Orange")+
  labs(title = "TotalSteps vs Total Time In Bed")+theme(plot.title=element_text(hjust=0.5))

```

#### There seems to be no real direct correlation between total steps taken through the day and how long you sleep. 
#### To help people wind down for the day, we can provide relaxation recommendations an hour before their average bedtime.

### Let's compare tracker distance with calories burned.
```{r}
ggplot(data=combined_data, aes(x=TrackerDistance, y=Calories))+geom_point(color="NavyBlue")+geom_smooth(color="Orange")+labs(title = "TrackerDistance vs Calories Burned")+theme(plot.title=element_text(hjust=0.5))
```

#### The longer the distance tracked is, the more calories are burned.




### Recommendations based off analysis
#### We seem to have little to no data on water intake and weight tracking. I would recommend that the company develop a smart scale that communicates with the app.
#### I would call it "Feather" to coincide with having a positive attitude about checking your weight on a daily basis. 
#### We could offer a 10% discount code to new app subscribers for the Spring water bottle or the "Feather" smart scale.
#### We could also offer a 20% discount code to new app subscribers who want to buy both the Spring water bottle and the "Feather" smart scale together.


#### This is my very first analysis I've ever done. I am a complete novice. Any recommendations and constructive criticism is welcome. Please be kind.


 